Mortalidad en áreas menores de la Región Pampeana (2009-2011)

### 2020-04-26 analysis/Paper_draft_ENG.tex

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={A Methodological Approach Comparison of Mortality Estimates for Small Areas: the Pampean Region in Argentina (2009-2011)},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\title{A Methodological Approach Comparison of Mortality Estimates for Small
Areas: the Pampean Region in Argentina (2009-2011)}
\author{Nicolás Sacco;\footnote{Penn State,
  \href{mailto:nsacco@psu.edu}{\nolinkurl{nsacco@psu.edu}}} Iván
Williams;\footnote{Universidad Nacional de Luján,
  \href{mailto:ivanwilliams1985@gmail.com}{\nolinkurl{ivanwilliams1985@gmail.com}}}
Bernardo L. Queiroz\footnote{Cedeplar-UFMG,
  \href{mailto:lanza@cedeplar.ufmg.br}{\nolinkurl{lanza@cedeplar.ufmg.br}}}}
\date{Aug, 2019}

@@

Life expectancy at birht was calculated based on a Bayesian approach, a
relational life-table method, smoothing spatial data `borrowing
strength' from neighboring areas, and third a indirect demographic
approach. Estimates are compared to calculate q0 rates from full death
registers and then related to spatial areas, using a simple
regionalization analysis.

@@

\hypertarget{introduction}{%
\subsection{\texorpdfstring{\textbf{Introduction}}{Introduction}}\label{introduction}}

In Latin America and the Caribbean, the demand for epidemiological (and
mortality, specifically) estimates of the sub-national heterogeneity is
growing, both as a tool for the application of different development
plans and for the allocation of resources. The experience in Latin
America is led by Brazil, where there is already a sustained
methodological development Freire
(\protect\hyperlink{ref-FreireEtAl2015}{2015}). In Argentina,
experiences were found on estimates of the infant mortality trend
(Torcida, Vega, and Velázquez
(\protect\hyperlink{ref-torcida2008}{2008})), and the communes of the
Autonomous City of Buenos Aires (Grushka
(\protect\hyperlink{ref-Grushka2013}{2013})). At the same time, more
studies are focused in developing demographic tools for measuring
individual heterogeneity in a population (Vaupel and Missov
(\protect\hyperlink{ref-Vaupel_Missov_2013}{2013}); Raalte, Sasson, and
Martikainen
(\protect\hyperlink{ref-vanRaalte_Sasson_Martikainen_2018}{2018})). In
this sense, only considering life table information, there could be
multiple layers of inequality analysis just by grouping by
administrative levels.

The objective of this work can be enunciated following these progressive
steps:\\
+ Estimate the level of mortality of the departments during the period
2009-2011. With the criteria of starting with the population with bigger
share to explore methodological approaches, the Pampean region was
selected.\\
+ Estimate the heterogeneity between and within departments.\\
+ Test the hypothesis of the negative relationship between internal
heterogeneity and the level of mortality at the province level.

In this first work, the step 1 is done.\footnote{All the supplemental
  material, code and results can be found in
  \url{https://github.com/nsacco/mortalidad_Argentina}. It´s work in
  progress}

@@

{\centering \includegraphics[width=0.7\linewidth]{plots/AjusteFeos}

@@

{\centering \includegraphics[width=0.7\linewidth]{plots/dptosBsAs}

@@

\leavevmode\hypertarget{ref-FreireEtAl2015}{}%
Freire, Queiroz, F. H. M. d. A. 2015. ``Mortality Estimates and
Construction of Life Tables for Small Areas in Brazil, 2010.''

@@

\leavevmode\hypertarget{ref-vanRaalte_Sasson_Martikainen_2018}{}%
Raalte, Alyson A. van, Isaac Sasson, and Pekka Martikainen. 2018. ``The
Case for Monitoring Life-Span Inequality.'' \emph{Science} 362 (6418):
1002--4. \url{https://doi.org/10.1126/science.aau5811}.

@@

\leavevmode\hypertarget{ref-torcida2008}{}%
Torcida, Sebastián, Andrea L Vega, and Guillermo A Velázquez. 2008.
``Análisis de La Evolución de La Tasa de Mortalidad Infantil En Los
Departamentos de Argentina. 1994-20.''

\leavevmode\hypertarget{ref-Vaupel_Missov_2013}{}%
Vaupel, James W, and Trifon I Missov. 2013. ``Unobserved Population
Heterogeneity: A Review of Formal Relationships.'' \emph{Demographic
Research} 31. Demographic Research: 659--86.
\url{https://www.demographic-research.org/volumes/vol31/22/default.htm}.

>>>>

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={A Methodological Approach Comparison to Estimate Mortality in Small Areas: the Pampean Region in Argentina (2009-2011)},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\title{A Methodological Approach Comparison to Estimate Mortality in Small
Areas: the Pampean Region in Argentina (2009-2011)}
\author{Nicolás Sacco;\footnote{Penn State,
  \href{mailto:nsacco@psu.edu}{\nolinkurl{nsacco@psu.edu}}} Iván
Williams;\footnote{Universidad Nacional de Luján,
  \href{mailto:ivanwilliams1985@gmail.com}{\nolinkurl{ivanwilliams1985@gmail.com}}}
Bernardo L. Queiroz\footnote{Cedeplar-UFMG,
  \href{mailto:lanza@cedeplar.ufmg.br}{\nolinkurl{lanza@cedeplar.ufmg.br}}}}
\date{Aug, 2019}

@@

Life expectancy at birht was calculated based on a Bayesian approach, a
relational life-table method, smoothing spatial data `borrowing
strength' from neighboring areas, and third a indirect demographic
approach. Estimates are compared to calculate \(q_0\) rates from full
death registers and then related to spatial areas, using a simple
regionalization analysis.

@@

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

@@

{\centering \includegraphics[width=0.7\linewidth]{analysis/plots/AjusteFeos}

@@

{\centering \includegraphics[width=0.7\linewidth]{analysis/plots/dptosBsAs}

@@

@@

@@

### 2020-04-26 analysis/Paper_draft_ESP.Rmd

```{r SinDEP, echo = F, results='asis', error=F, warning=F}
library(kableExtra)
kable(SinDEP, digits = 2, caption = 'Pronvincias según Departamento desconocido',
      col.names = c("Province", 'Unknown %')) %>% #'$Ls(IC95)$',
      add_footnote(c("Source: elaboración propia"))
```

```{r UnkSexAge, echo = F, results='asis', error=F, warning=F}
kable(UnkSexAge, digits = 2, caption = 'Departamentos con información desconocida', row.names = F) %>%
      add_footnote(c("Source: elaboración propia"))
```

```{r def_tardias, echo = F, results='asis', error=F, warning=F}
kable(def_tardias, caption='Distribución de defunciones registradas y ocurridas') %>%
      add_footnote(c("The year register is in rows, and ocurred in columns. Source: elaboración propia"))  %>%
      add_header_above(c(" " = 1, "Ocurred year" = 6))
```

```{r Dif_e0_INDEC, echo = F, results='asis', error=F, warning=F}
library(kableExtra)
kable(Dif_e0_INDEC, digits = 2, caption='Diferencias entre la esperanza de vida calculada con datos no ajustados y estimaciones oficiales',
      col.names = c('Province', "Own", "INDEC", 'Realtive Diference')) %>%
      kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
      full_width = F, position = "center")%>%
      add_footnote(c("Source: elaboración propia"))
```

```{r table_e0, echo = F, results='asis', error=F, warning=F}
library(kableExtra)
kable(table_e0, digits = 2, caption='Estimación de la esperanza de vida al nacer. Córdoba 2008-2010',
      col.names = c('Province', "Department", "Mean", 'p97.5', 'p2.5')) %>%
      kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
      full_width = F, position = "center")%>%
      add_footnote(c("Source: elaboración propia"))
```

>>>>

```{r SinDEP, echo = F, results='asis', error=F, warning=F}
library(kableExtra)
kable(SinDEP, digits = 2, caption = 'Pronvincias según Departamento desconocido',
      col.names = c("Provincia", 'Desconocido %')) %>% #'$Ls(IC95)$',
      add_footnote(c("Fuente: elaboración propia"))
```

```{r UnkSexAge, echo = F, results='asis', error=F, warning=F}
kable(UnkSexAge, digits = 2, caption = 'Departamentos con información desconocida', row.names = F) %>%
      add_footnote(c("Fuente: elaboración propia"))
```

```{r def_tardias, echo = F, results='asis', error=F, warning=F}
kable(def_tardias, caption='Distribución de defunciones registradas y ocurridas') %>%
      add_footnote(c("El año de registro está en las filas y el de ocurrencia en las columnas. Fuente: elaboración propia"))  %>%
      add_header_above(c(" " = 1, "Año de ocurrencia" = 6))
```

```{r Dif_e0_INDEC, echo = F, results='asis', error=F, warning=F}
library(kableExtra)
kable(Dif_e0_INDEC, digits = 2, caption='Diferencias entre la esperanza de vida calculada con datos no ajustados y estimaciones oficiales',
      col.names = c('Provincia', "Propia", "INDEC", 'Diferencia relativa')) %>%
      kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
      full_width = F, position = "center")%>%
      add_footnote(c("Fuente: elaboración propia"))
```

```{r table_e0, echo = F, results='asis', error=F, warning=F}
library(kableExtra)
kable(table_e0, digits = 2, caption='Estimación de la esperanza de vida al nacer. Córdoba 2008-2010',
      col.names = c('Province', "Department", "Mean", 'p97.5', 'p2.5')) %>%
      kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
      full_width = F, position = "center")%>%
      add_footnote(c("Fuente: elaboración propia"))
```
